---
title: "QMM Final Project"
author: "Sharik Baig, Sai Venkata, Rajiv Reddy"
date: "12/10/2022"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Constraints

The Task is to assign 12 students to 4 groups with not more than 3 students in each group.

Hence we have considered 3 major factors that is very important for best group distribution.

Student's GPA - Considering the example that GPA 3 is a good GPA hence we have taken sum of GPA for each group should be >= 9.0
Student's Attendance - Considering the example that attendance should be atleast 60% hence we have taken sum of attendance for each group should be >= 180.0
Student's grade from Discussion Post - Considering the example that avg grade from discussion post should be atleast 70% hence we have taken sum of grade from discussion post for each group should be >= 210.0

```{r students}
library(lpSolveAPI)
StuGroup <- read.lp("C:/Users/shari/OneDrive/Desktop/Business Analytics/QMM/QMM_Project_Group6/Group.lp")
StuGroup
solve(StuGroup)
get.objective(StuGroup)
get.variables(StuGroup)
get.constraints(StuGroup)
```

## Group Allocation

Looking at the output below is the group allocation

Group1 - Student3, Student6, Student9.
Group2 - Student5, Student8, Student10.
Group3 - Student4, Student7, Student1.
Group4 - Student2, Student11, Student12.

