---
title: "Assignment "
output: html_document
---
                         Module 6 - The Transportation Model
Heart Start produces automated external defibrillators (AEDs) in each of two different plants (A and B). The unit production costs and monthly production capacity of the two plants are indicated in the table below. The AEDs are sold through three wholesalers. The shipping cost from each plant to the warehouse of each wholesaler along with the monthly demand from each wholesaler are also indicated in the table. How many AEDs should be produced in each plant, and how should they be distributed to each of the three wholesaler warehouses so as to minimize the combined cost of production and shipping?


1. Formulate and solve this transportation problem using R
2. Formulate the dual of this transportation problem
3. Make an economic interpretation of the dual

```{r}
tab <- matrix(c(22,14,30,0,16,20,24,0), ncol = 4,byrow = TRUE)
# Set customers and suppliers' names
colnames(tab) <- c("Warehouse 1", "Warehouse 2", "Warehouse 3","Dummy")
rownames(tab) <- c("Plant A", "Plant B")
tab
```

The above transportation problem can be formulated in the LP format as below:

Min TC = 22x11 + 14x12 + 30x13 
        +16x21 + 20x22 + 24x23
        
        
Subject to          
Production Capacity        
       x11 + x12 + x13  ≤ 100
       x21 + x22 + x23  ≤ 120

Demand constrain
      x11 + x21  ≥ 80
      x21 + x22  ≥ 60
      x21 + x22  ≥ 70

Non-negativity of the variables
            xij ≥ 0
          
where
            i = 1,2,3
and 
            j = 1,2
 


This transportation problem is unbalanced one (demand is not equal to supply), that is demand is less than supply by 10,
so I create a dummy variable in column 4 with transportation cost zero and demand 10

```{r}
library(lpSolve)
# Set up cost matrix
costs <- matrix(c(22,14,30,0,16,20,24,0), ncol = 4,byrow = TRUE)
# Set customers and suppliers' names
colnames(costs) <- c("Warehouse 1", "Warehouse 2", "Warehouse 3","Dummy")
rownames(costs) <- c("Plant A", "Plant B")
costs
```


```{r}
# Set up constraint signs and right-hand sides (supply side)
row.signs <- rep("<=", 2)
row.rhs <- c(100,120)
#Demand (sinks) side constraints
col.signs <- rep(">=", 4)
col.rhs <- c(80,60,70,10)
```



```{r}
# Run
lptrans <- lp.transport(costs, "min", row.signs, row.rhs, col.signs, col.rhs)
#Values of all 9 variables
lptrans$solution
```


```{r}
# Value of the objective function
lptrans$objval
```

When solved the transportation problem, I got the values of the variables as
x11 = 30
x12 = 50
x21 = 60
x32 = 70


The minimized value of the transportation cost is 3980
