---
title: "Assignment 6"
output:
  pdf_document: default
  html_document: default
date: "2022-11-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(lpSolveAPI)
```
AP shipping service worker scheduling 

```{r}
# create an lp object with 0 constraints and 7 decision variables
lprec <- make.lp(0, 7)

# Making an objective function. The default problem type is a minimization problem.
set.objfn(lprec, c(775,800,800,800,800,775,750))

# Adding the constraints
add.constraint(lprec, c(0, 1, 1, 1, 1 ,1, 0), ">=", 18)
add.constraint(lprec, c(0, 0, 1, 1, 1 ,1, 1), ">=", 27)
add.constraint(lprec, c(1, 0, 0, 1, 1 ,1, 1), ">=", 22)
add.constraint(lprec, c(1, 1, 0, 0, 1 ,1, 1), ">=", 26)
add.constraint(lprec, c(1, 1, 1, 0, 0 ,1, 1), ">=", 25)
add.constraint(lprec, c(1, 1, 1, 1, 0 ,0, 1), ">=", 21)
add.constraint(lprec, c(1, 1, 1, 1, 1 ,0, 0), ">=", 19)

# Set bounds for variables explicitly.
set.bounds(lprec, lower = c(0, 0, 0, 0, 0, 0, 0), columns = c(1, 2,3,4,5,6,7))

# Making the decision variables as Integer 
set.type(lprec,1:7,"integer")

# Nameing the decision variables (column) and constraints (rows)
lp.rownames <- c("Sunday", "Monday", "Tuesday", "Wednesady", "Thursday", "Friday", "Saturday")

#Rows represents the day shift starts 
lp.colnames <- c("Shift1", "Shift2", "Shift3", "Shift4", "Shift5", "Shift6", "Shift7")
dimnames(lprec) <- list(lp.rownames, lp.colnames)

# Examine the linear program object to ensure that it is correct.
lprec

# Writing the model to a file
write.lp(lprec, filename = "APShipping.lp", type = "lp")

```
# In the table below, we've estimated how many people will be needed each day for a week.
```{r}
Workersrequired_Daywise<-matrix(c("Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday",18,27,22,26,25,21,19),ncol=2,byrow=)
colnames(Workersrequired_Daywise)<-c("Days_in_aweek","Number_of_Required_workers")

as.table(Workersrequired_Daywise)
```

```{r}
Daysoff_and_Wages<-matrix(c(1,2,3,4,5,6,7,"Sunday and monday","Monday and Tuesday","Tuesday and wednesday","wednesday and Thursday","Thursday and friday","Friday and Saturday","Saturday and Sunday","$775","$800","$800","$800","$800","$775","$750"),ncol = 3,byrow = F)
colnames(Daysoff_and_Wages)<-c("Shifts","Days_offs","Wages")
as.table(Daysoff_and_Wages)
```

```{r}

# Solving it as a LP model
solve(lprec)

# Objective function value:
get.objective(lprec)

# Total wage cost :25675

# Optimal decision variable values:
get.variables(lprec)

# Values of the Constraints 
get.constraints(lprec)

```
#The overall cost to the company to ensure that total pay expenses are kept to a minimum and there are enough workers available for work each day is $ 25675
#Shift 1 No. of workers: 18
#Shift 2 No. of workers: 27
#Shift 3 No. of workers: 25
#Shift 4 No. of workers: 26
#Shift 5 No. of workers: 27
#Shift 6 No. of workers: 23
#Shift 7 No. of workers: 19

